---
title: "Data Visualization Assignment"
output: html_notebook
---

Amy's data visualizaation assignment for PNB3EE3

# Loading in data

## Installing required packages

```{r}
list.of.packages <- c("tidyverse", "ggpubr", "ez", "corrplot", "car", "stats","PerformanceAnalytics", "afex", "emmeans", "mlma", "ggplot2", "ggExtra", "psych", "ggpubr", "here", "gridExtra", "ggridges","hrbrthemes")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages) 
```

##Loading in csv file Loading in csv file Load in csv file from <https://perception-lab-pnb3ee3.github.io/courseBook/assignment_dataVis.html>

```{r}
here::i_am("dataVisualisation_analyses.Rmd")
library(here)
data <- read.csv(here("assignment_dataVis_dataset.csv")) 
fig_path <- here("figures/") # path to save generated figures to
op = 0 # flag if want to save output plots to files (1) or plot inline (0)
```

# Summary Statistics

*preliminary inspection of the file and properties of the data*

```{r}
print(summary(data))
head(data)
```

## *Sorting data by conditions*

```{r}
df_list <- split(data, data$condition)

df_condition1 <- df_list[["condition_1"]]
df_condition2 <- df_list[["condition_2"]]
df_condition3 <- df_list[["condition_3"]]
df_condition4 <- df_list[["condition_4"]]

```

## Condition 1 statistics

```{r}
describe(df_condition1)
```

## Condition 2

```{r}
describe(df_condition2)
```

## Condition 3 statistics

```{r}
describe(df_condition3)
```

## Condition 4 statistics

```{r}
describe(df_condition4)
```

# Data Visualization

## Condition 1

```{r}
library(ggplot2)
library(hrbrthemes)
condition1_plot <- ggplot(df_condition1, aes(x=x, y=y)) +
  geom_point() +
  geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) +
  theme_ipsum()
condition1_plot
```

## Condition 2

```{r}
condition2_plot <- ggplot(df_condition2, aes(x=x, y=y)) +
  geom_point() +
  geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) +
  theme_ipsum()
condition2_plot
```

## Condition 3

```{r}
condition3_plot <- ggplot(df_condition3, aes(x=x, y=y)) +
  geom_point() +
  geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) +
  theme_ipsum()
condition3_plot
```

## Condition 4

```{r}
condition4_plot <- ggplot(df_condition4, aes(x=x, y=y)) +
  geom_point() +
  geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) +
  theme_ipsum()
condition4_plot
```

# Reflection

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
