---
title: "Data Visualization Assignment"
output: html_notebook
---

Amy's data visualizationassignment for PNB3EE3

# Loading in data

## Installing required packages

```{r, results='hide'}
list.of.packages <- c("tidyverse", "ggpubr", "ez", "corrplot", "car", "stats","PerformanceAnalytics", "afex", "emmeans", "mlma", "ggplot2", "ggExtra", "psych", "ggpubr", "here", "gridExtra", "ggridges","hrbrthemes","psych","ggpmisc")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages) 
library(psych)
library(ggplot2)
library(hrbrthemes)
library(ggpmisc)
```

## Loading in csv file 
Loading in csv file Load in csv file from <https://perception-lab-pnb3ee3.github.io/courseBook/assignment_dataVis.html>

```{r,echo=TRUE, results='hide'}
here::i_am("dataVisualisation_analyses.Rmd")
library(here)
data <- read.csv(here("assignment_dataVis_dataset.csv")) 
op = 0 # saves plot inline (0) instead of in a file (1)
```

# Summary Statistics

*Preliminary inspection of the file and properties of the data*

```{r}
print(summary(data))
head(data)
```

## *Sorting data by conditions*

```{r, echo=TRUE, results='show'}
df_list <- split(data, data$condition)

df_condition1 <- df_list[["condition_1"]]
df_condition2 <- df_list[["condition_2"]]
df_condition3 <- df_list[["condition_3"]]
df_condition4 <- df_list[["condition_4"]]

```

## Condition 1 statistics

```{r, echo=TRUE, results='show'}
describe(df_condition1)
```

## Condition 2 statistics

```{r, echo=TRUE, results='show'}
describe(df_condition2)
```

## Condition 3 statistics

```{r, echo=TRUE, results='show'}
describe(df_condition3)
```

## Condition 4 statistics

```{r, echo=TRUE, results='show'}
describe(df_condition4)
```

# Data Visualization

## Condition 1

```{r, echo=TRUE, results='show'}
condition1_plot <- ggplot(df_condition1, aes(x=x, y=y)) +
  geom_point(size=1, alpha=0.6) +
  geom_smooth(method=lm , color="red", fill="#6699ff", se=TRUE) +
    stat_poly_eq(aes(label = ..eq.label..), 
    formula = y ~ x, 
    label.x = max(df_condition1$x) * 0.5,  # Position near the right
    label.y = max(df_condition1$y) * 0.9   # Position near the top
  )+
  theme_ipsum() +
   labs(
    title = "Condition 1",
    subtitle = "Linear regression with confidence interval",
    x = "X Variable Label",
    y = "Y Variable Label"
    )+
  xlim(0,100)+ylim(0,100)
condition1_plot
```

## Condition 2

```{r, echo=TRUE, results='show'}
condition2_plot <- ggplot(df_condition2, aes(x=x, y=y)) +
  geom_point(size=1, alpha=0.6) +
  geom_smooth(method=lm , color="red", fill="#6699ff", se=TRUE) +
    stat_poly_eq(aes(label = ..eq.label..), 
    formula = y ~ x, 
    label.x = max(df_condition2$x) * 0.5, 
    label.y = max(df_condition2$y) * 0.9  
  )+
  theme_ipsum() +
   labs(
    title = "Condition 2",
    subtitle = "Linear regression with confidence interval",
    x = "X Variable Label",
    y = "Y Variable Label"
    )+
  xlim(0,100)+ylim(0,100)
condition2_plot
```

## Condition 3

```{r, echo=TRUE, results='show'}
condition3_plot <- ggplot(df_condition3, aes(x=x, y=y)) +
  geom_point(size=1, alpha=0.6) +
  geom_smooth(method=lm , color="red", fill="#6699ff", se=TRUE) +
    stat_poly_eq(aes(label = ..eq.label..), 
    formula = y ~ x, 
    label.x = max(df_condition3$x) * 0.5, 
    label.y = max(df_condition3$y) * 0.9 
  )+
  theme_ipsum() +
   labs(
    title = "Condition 3",
    subtitle = "Linear regression with confidence interval",
    x = "X Variable Label",
    y = "Y Variable Label"
    )+
  xlim(0,100)+ylim(0,100)
condition3_plot
```

## Condition 4

```{r, echo=TRUE, results='show'}
condition4_plot <- ggplot(df_condition4, aes(x=x, y=y)) +
  geom_point(size=1, alpha=0.6) +
  geom_smooth(method=lm , color="red", fill="#6699ff", se=TRUE) +
    stat_poly_eq(aes(label = ..eq.label..), 
    formula = y ~ x, 
    label.x = max(df_condition4$x) * 0.5, 
    label.y = max(df_condition4$y) * 0.9   
  )+
  theme_ipsum() +
   labs(
    title = "Condition 4",
    subtitle = "Linear regression with confidence interval",
    x = "X Variable Label",
    y = "Y Variable Label"
    )+
  xlim(0,100)+ylim(0,100)
condition4_plot
```

# Reflection
**What can be concluded from the dataset?**

 - Even though the summary statistics, regression, and correlation coefficients are very similar (or even exactly the same) across the four conditions, the data distributions could be very different
- And this dataset was probably created by someone artistic

**What did you find most challenging about this assignment?**

I found it surprisingly challenging to get R to isolate each condition and operate on each condition separately. Mostly because I didn't know what kind of functions are available in R and what they do exactly. I first tried to convert the csv from long to wide, (and some other attempts I've forgotten) but then I couldn't separate x and y when doing summary statistics. 
Eventually I figured out that i could define each condition as a separate dataframe and operate on them individually. 

**What did you learn in completing this assignment? What are you still curious about?**

 - Learned that R notebook is very powerful and fun
- And how to find coding help with the Internet (and that it's okay to not know all they syntax of every language you use)


- Curious about how to deal with more complex types of data using R, beyond numeric values
- And maybe how the data analysis tools we have might limit our experimental design process
  - e.g. working with language, text, video, doing thematic analysis, etc. 
  
**Why is it important to visualize data?**

 - It's a MUCH easier way to communicate and share the results compared to text, and helps widen the audience of your research
- Helps you inspect the shape and distribution of the data and detect any anomalous patterns or outliers
- It's fun and it makes your report/paper more visually pleasing and readable :)